---
title: "Untitled"
author: "Roy Thomas"
date: "2025-09-07"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Q1. Gapminder bar chart with reordered factors
**Task.** Using `gapminder`:
- Filter to the year **2007**.
- Compute **average life expectancy** by **continent**.
- Make a **bar chart** with continents reordered by average life expectancy.
- Format the y-axis to show **whole numbers** only and add clear labels.


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
library(tidyverse)   # ggplot2, dplyr, readr, tibble, tidyr, stringr, forcats
library(ggthemes)    # scale_color_colorblind()
library(gapminder)   # gapminder dataset
library(ggrepel)     # geom_text_repel (for cleaner labels)
library(scales)      # nice axis labels/breaks
```

```{r q1-solution}
gap_2007 <- gapminder |>
  filter(year == 2007) |>
  group_by(continent) |>
  summarize(avg_lifeExp = mean(lifeExp, na.rm = TRUE), .groups = "drop")

# When you add .groups = "drop" inside summarize(), you’re explicitly telling dplyr:
# "Remove the grouping after summarizing, give me back a plain tibble.”
# Otherwise, if you didn't have the "drop" line, then it may affect future operations

ggplot(gap_2007, aes(x = fct_reorder(continent, avg_lifeExp), y = avg_lifeExp)) +
  geom_col(fill = "steelblue") +
  scale_y_continuous(labels = number_format(accuracy = 1),
                     limits = c(0, max(gap_2007$avg_lifeExp) * 1.1),
                     expand = expansion(mult = c(0, 0.05))) +
  labs(
    x = NULL,
    y = "Average Life Expectancy (years)",
    title = "Average Life Expectancy by Continent (2007)",
    subtitle = "Gapminder"
  ) 
  #coord_cartesian(ylim = c(0, max(gap_2007$avg_lifeExp))) 
  #theme_minimal(base_size = 12)
```



## Q2. Scatterplot with logarithmic scale and labels
**Task.** Using built-in `mtcars`:
- Scatterplot of `wt` (x) vs `mpg` (y).
- Add car names as **labels**.
- Transform the x-axis to **log10** scale and relabel it “Car Weight (log scale)”.
- Keep the plot readable.



```{r q2-solution}
#If you look at the mtcars dataset, you will see that the ROW NAMES
#are the car names.  If you want to see the car names, as a COLUMN,
#we would need to run the rownnames_to_column command

#Take a look at mtcars
#mtcars

#Now, let us make the car names into a COLUMN
mtcars_tbl <- rownames_to_column(mtcars, var = "car")


ggplot(mtcars_tbl, aes(x = wt, y = mpg, label = car)) +
  geom_point() +
  #Setting min.segment.length = 0 forces all leader lines to be drawn, even if they’re tiny.
  geom_text_repel(min.segment.length = 0, size = 2) +
  scale_x_log10(name = "Car Weight (log scale)",
                labels = label_number(accuracy = 0.1)) +
  labs(y = "Miles per Gallon (mpg)",
       title = "Fuel Efficiency vs. Vehicle Weight",
       subtitle = "mtcars dataset (x-axis log10)") +
  theme_minimal(base_size = 12)

```

## Q3. Species scatterplot with regression lines
**Task.** Using `iris`:
- Scatterplot of `Sepal.Length` vs `Petal.Length`.
- Color points by `Species`.
- Add **separate regression lines** per species.
- Use a colorblind-friendly palette.


```{r q3-solution}
ggplot(iris, aes(x = Sepal.Length, y = Petal.Length, color = Species)) +
  geom_point(alpha = 0.8) +
  #by default, ggplot2 adds a shaded ribbon around the line, 
  #representing the 95% confidence interval for the fitted values.
  #se = FALSE, turns that shading off, so you see just the line.
  geom_smooth(method = "lm", se = FALSE) +
  scale_color_colorblind() +
  labs(
    x = "Sepal Length (cm)",
    y = "Petal Length (cm)",
    title = "Iris: Petal vs Sepal Length by Species",
    color = "Species"
  ) +
  theme_minimal(base_size = 12)

```





## Q4. Distribution plots: histogram and density (diamonds)
**Task.** Using `ggplot2::diamonds`:
- Plot a **histogram** of `price` with `binwidth = 500`.
- On a separate plot, show a **density curve** of `price` (use a visible outline color).
- Briefly compare what each reveals about **skewness**.


```{r q4-solution}
# Histogram
ggplot(diamonds, aes(x = price)) +
  #boundary = 0  : Controls where the bins start.  
  #closed = "left" : Determines which side of the bin interval is inclusive:
  geom_histogram(binwidth = 500, boundary = 0, closed = "left", fill = "grey70") +
  labs(
    x = "Price (USD)",
    y = "Count",
    title = "Diamonds: Price Distribution (Histogram, binwidth = 500)"
  ) +
  theme_minimal(base_size = 12)

# Density
ggplot(diamonds, aes(x = price)) +
  geom_density(linewidth = 1) +
  labs(
    x = "Price (USD)",
    y = "Density",
    title = "Diamonds: Price Distribution (Density)"
  ) +
  theme_minimal(base_size = 12)

```


**Instructor note (talk track):**  
Both plots show a **right-skewed** distribution. The histogram emphasizes counts per bin and can reveal local spikes caused by pricing granularity; the density smooths these into an overall shape, making the **skewness** more visually apparent.

## Q5. Faceted boxplots (mpg)
**Task.** Using `mpg`:
- Boxplot of `hwy` by `class`.
- **Facet** by `drv` (f = front, r = rear, 4 = 4-wheel).
- Rotate x-axis labels for readability.


```{r q5-solution}
ggplot(mpg, aes(x = class, y = hwy)) +
  #This line below makes the outlier partially transparent
  geom_boxplot(outlier.alpha = 0.5) +
  facet_wrap(~ drv) +
  labs(
    x = "Vehicle Class",
    y = "Highway MPG",
    title = "Highway Fuel Economy by Class, Faceted by Drive Type"
  ) +
  theme_minimal(base_size = 12) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

```


