---
title: "In-Class Exercise: ggplot Practice"
author: "Your Name"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
library(ggplot2)
library(dplyr)
library(forcats)
library(palmerpenguins)   # for penguins dataset
```

---

## Exercise 1: Scatterplot with Regression Line  
Dataset: `mpg` (built into ggplot2)  

> Create a scatterplot of engine displacement (`displ`) vs. highway miles per gallon (`hwy`).  
> - Add a **linear regression line** (`geom_smooth(method="lm")`).  
> - Color the points by car class (`class`).  
> What relationship do you see?  


```{r}





```

---

## Exercise 2: Bar Chart of Counts  
Dataset: `penguins` (palmerpenguins package)  

> Plot a bar chart showing the number of penguins by `species`.  
> - Reorder the bars so that the most common species comes LAST  

```{r}






```

---

## Exercise 3: Line Chart of Trends  
Dataset: `economics` (built into ggplot2)  

> Plot the unemployment (`unemploy`) over time (`date`).  
> - Use a line chart.  
> - Add a smoother with `geom_smooth(method="loess")`.  
> What long-term trend do you observe?  

```{r}




```

---

## Exercise 4: Boxplot with Jittered Points  
Dataset: `diamonds` (built into ggplot2)  

> Create a boxplot of diamond price (`price`) by diamond cut (`cut`).  
> This would mean that price is y, and cut is x
> - Overlay the raw points using `geom_jitter(alpha=0.1)`.  
> - Which cut tends to have the highest median price?  

```{r}





```

---

## Exercise 5: Stacked Bar Plot with Fill  
Dataset: `mpg` (built into ggplot2)  

> Plot a stacked bar chart of number of cars by `class`, filled by `drv` (drive type: f = front-wheel, r = rear-wheel, 4 = 4-wheel).  
> - Use `geom_bar(position="fill")` so the y-axis shows proportions.  
> - Which car class has the highest proportion of 4-wheel drive?  

```{r}





```
