---
title: "BUA 477"
description: Visual Analytics
site: distill::distill_website
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

# Learn more about creating websites with Distill at:
# https://rstudio.github.io/distill/website.html

# Learn more about publishing to GitHub Pages at:
# https://rstudio.github.io/distill/publish_website.html#github-pages

```

This is the home page for BUA 477 - Visual Analytics

Instructor: Roy Thomas
Meeting times: Tu/Th - 11am to 12:15pm 

To run any of the materials locally on your own machine, you will need the following:

- A recent version of R, [download from here.](https://cloud.r-project.org/)
- A recent version of RStudio, [download from here.](https://rstudio.com/products/rstudio/download/#download)
- The following R packages:    
  broom, cluster, colorspace, cowplot, distill, gapminder, GGally, gganimate,  ggiraph, ggdendro, ggdist, ggforce, ggplot2movies, ggrepel, ggridges, ggthemes, gifski, glue, knitr, learnr, naniar, margins, MASS, Matrix, nycflights13, palmerpenguins, patchwork, rgdal, rmarkdown, rnaturalearth, rnaturalearthhires, sf, shinyjs, tidyverse, transformr, umap, xaringan


You can install all required R packages at once by running the following code in the R command line:  (HOWEVER, WE WILL NOT INSTALL THEM ALL TOGETHER, RIGHT NOW. INSTEAD, WE WILL ONLY INSTALL THE PACKAGES WE NEED.  SO IGNORE THESE INSTALLATIONS, UNTIL TOLD LATER!!)

```{r eval = FALSE, echo = TRUE}
# first run this command:
install.packages(
  c(
    "broom", "cluster", "colorspace", "cowplot", "distill", "gapminder", 
    "GGally", "gganimate", "ggiraph", "ggdendro", "ggdist", "ggforce",
    "ggplot2movies", "ggrepel", "ggridges", "ggthemes", "gifski", "glue",
    "knitr", "learnr", "naniar", "margins", "MASS", "Matrix",
    "nycflights13", "palmerpenguins", "patchwork", "rgdal", "rmarkdown",
    "rnaturalearth", "sf", "shinyjs", "tidyverse", "transformr", "umap",
    "xaringan"
  )
)

#HOLD OFF RUNNING THIS COMMAND, AS WELL
# then run this command: 
install.packages(
  "rnaturalearthhires", repos = "https://packages.ropensci.org", type = "source"
)
```

## Reuse {.appendix}

Text and figures are licensed under Creative Commons Attribution [CC BY 4.0](https://creativecommons.org/licenses/by/4.0/). Any computer code (R, HTML, CSS, etc.) in slides and worksheets, including in slide and worksheet sources, is also licensed under [MIT](LICENSE.html). Note that figures in slides may be pulled in from external sources and may be licensed under different terms. For such images, image credits are available in the slide notes, accessible via pressing the letter 'p'.
Some materials are: Copyright (c) 2021 Claus O. Wilke


