---
title: "Untitled"
author: "Roy Thomas"
date: "2025-09-18"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

---

title: "Animation Plots – 5-Problem Exercise"
author: "Your Name"
date: "`r Sys.Date()`"
output:
html\_document:
df\_print: paged
toc: true
toc\_float: true
----------------

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

> **Instructions**
>
> * This set has **5 problems**. Each mirrors an animated example from class and adds **one small extension** to research.
> * Use built-in datasets where possible (`gapminder`, `economics`, `storms`, etc.).
> * Where you see **Stretch**, look up the function/argument and implement it.
> * If you need a hint, see the **commented Instructor solution** chunk below each problem (keep it commented for the student version).

## Packages

```{r}
# Install (first time only) then load
# install.packages(c("ggplot2","dplyr","gganimate","gapminder","ggrepel","maps","rnaturalearth","rnaturalearthdata","scales"))

library(ggplot2)
library(dplyr)
library(gganimate)
library(gapminder)
library(ggrepel)
library(scales)
# for Problem 5 (map background)
library(rnaturalearth)
library(rnaturalearthdata)
```

---

## Problem 1 — Animated Scatter (gapminder over time)

Create an **animated scatterplot** of **life expectancy** vs **GDP per capita** from `gapminder`. Animate over `year`.

**Requirements**:

* `x = gdpPercap` (log10 scale), `y = lifeExp`, `size = pop`, `color = continent`.
* Use `transition_time(year)` and show the current year in the title using `{frame_time}`.
* Keep points from previous years faintly visible so motion is easier to track.

**Stretch** (research & add):

* Add a subtle trailing effect with `shadow_wake(wake = 0.1)` **or** `shadow_mark(alpha = 0.2)`.
* Use `view_follow()` so axes track the data without popping.

```{r}
# Your code here

 p1 <- gapminder %>%
   ggplot(aes(gdpPercap, lifeExp, size = pop, color = continent)) +
   geom_point(alpha = 0.8) +
   scale_x_log10(labels = label_number_si()) +
   scale_size_area(max_size = 12) +
   labs(title = "Year: {frame_time}", x = "GDP per capita (log)", y = "Life expectancy") +
   theme_minimal() +
   transition_time(year) +
   shadow_wake(wake = 0.1) +
   view_follow()

# animate(p1, fps = 20, width = 800, height = 500)


```

```{r, eval=FALSE}
# ---- Instructor solution (commented) ----
# p1 <- gapminder %>%
#   ggplot(aes(gdpPercap, lifeExp, size = pop, color = continent)) +
#   geom_point(alpha = 0.8) +
#   scale_x_log10(labels = label_number_si()) +
#   scale_size_area(max_size = 12) +
#   labs(title = "Year: {frame_time}", x = "GDP per capita (log)", y = "Life expectancy") +
#   theme_minimal() +
#   transition_time(year) +
#   shadow_wake(wake = 0.1) +
#   view_follow()
#
# # animate(p1, fps = 20, width = 800, height = 500)
```

---

## Problem 2 — Bar Chart Race (Top 10 GDP per capita)

Build a **bar chart race** showing the **top 10 countries by GDP per capita** for each year in `gapminder`.

**Requirements**:

* For each `year`, compute rank by `gdpPercap`, keep top 10.
* Horizontal bars: `y = rank` (reverse scale so 1 is at top), `x = gdpPercap`, `fill = country`, `group = country`.
* Left-label countries with `geom_text()` placed at a small negative x; widen limits with `xlim()` so labels fit.
* Animate with `transition_states(year)`; add `{closest_state}` to the title.

**Stretch** (research & add):

* Smooth transitions with `ease_aes('cubic-in-out')`.
* Balance slide vs. pause using `transition_states(year, transition_length = ..., state_length = ...)`.

```{r}
# Your code here
```

```{r, eval=FALSE}
# ---- Instructor solution (commented) ----
# top10 <- gapminder %>%
#   group_by(year) %>%
#   arrange(desc(gdpPercap), .by_group = TRUE) %>%
#   mutate(rank = row_number()) %>%
#   filter(rank <= 10) %>%
#   ungroup()
#
# p2 <- ggplot(top10, aes(gdpPercap, rank, group = country)) +
#   geom_col(aes(fill = country)) +
#   geom_text(aes(x = -500, y = rank, label = country), hjust = 1, size = 12/.pt) +
#   scale_y_reverse(breaks = 1:10) +
#   xlim(-4000, max(top10$gdpPercap) * 1.05) +
#   labs(title = "Year: {closest_state}", x = "GDP per capita", y = NULL) +
#   theme_minimal() + guides(fill = "none") +
#   transition_states(year, transition_length = 6, state_length = 3) +
#   ease_aes('cubic-in-out')
#
# # animate(p2, fps = 20, width = 900, height = 520)
```

---

## Problem 3 — Reveal Line with End Label (economics)

Create an **animated line** of `personal savings rate` from `ggplot2::economics` (column `psavert`). Use `transition_reveal(date)` so the line draws over time.

**Requirements**:

* Draw a line and small points; add an **end label** (the last value) using `geom_text_repel()` positioned slightly to the right.
* Ensure labels beyond the panel aren’t clipped: use `coord_cartesian(clip = "off")` and add extra right margin.

**Stretch** (research & add):

* Format y-axis as a **percentage** (divide by 100, use `scale_y_continuous(labels = scales::percent)`).
* Show the current date with `{frame_along}` in the subtitle.

```{r}
# Your code here
```

```{r, eval=FALSE}
# ---- Instructor solution (commented) ----
# df <- ggplot2::economics %>% mutate(pct = psavert/100)
# last_row <- tail(df, 1)
# p3 <- ggplot(df, aes(date, pct)) +
#   geom_line() +
#   geom_point(size = 0.8) +
#   geom_text_repel(
#     data = last_row,
#     aes(label = percent(pct, accuracy = 0.1)),
#     nudge_x = 100, direction = "y", hjust = 0
#   ) +
#   coord_cartesian(clip = "off") +
#   theme(plot.margin = margin(5, 60, 5, 5)) +
#   scale_y_continuous(labels = percent) +
#   labs(subtitle = "Date: {frame_along}") +
#   transition_reveal(date)
#
# # animate(p3, fps = 22, width = 900, height = 420)
```

---

## Problem 4 — Animated Rank Bump Chart (selected countries)

Pick **5–8 countries** from `gapminder` and plot how their **rank by life expectancy** changes over time (a **bump chart**). Animate the lines revealing over `year`.

**Requirements**:

* Compute `rank` by `lifeExp` within each `year`; keep only your selected countries.
* Plot `x = year`, `y = rank` (use `scale_y_reverse()` so rank 1 is top), `group = country`, `color = country`.
* Use `transition_reveal(year)` so each country’s rank trace draws over time.

**Stretch** (research & add):

* Add **end labels** to the right with `geom_text_repel()` and widen the right margin.
* Use a **discrete color palette** that is colorblind-friendly (`scale_color_brewer(palette = "Set2")` or `scale_color_viridis_d()`).

```{r}
# Your code here
```

```{r, eval=FALSE}
# ---- Instructor solution (commented) ----
# set.seed(1)
# selected <- sample(unique(gapminder$country), 6)
# ranks <- gapminder %>%
#   group_by(year) %>%
#   arrange(desc(lifeExp), .by_group = TRUE) %>%
#   mutate(rank = row_number()) %>%
#   ungroup() %>%
#   filter(country %in% selected)
#
# p4 <- ggplot(ranks, aes(year, rank, color = country)) +
#   geom_line(linewidth = 1) +
#   scale_y_reverse() +
#   labs(y = "Rank (life expectancy)", x = NULL) +
#   theme_minimal() +
#   transition_reveal(year)
#
# # Optional end labels
# # last_pts <- ranks %>% group_by(country) %>% slice_max(year, n = 1)
# # p4 <- p4 + ggrepel::geom_text_repel(
# #   data = last_pts,
# #   aes(label = country),
# #   nudge_x = 2, direction = "y", hjust = 0
# # ) + coord_cartesian(clip = "off") + theme(plot.margin = margin(5, 60, 5, 5))
#
# # animate(p4, fps = 20, width = 900, height = 480)
```

---

## Problem 5 — Animated Storm Track on a Map (`storms`)

Animate the track of a **single storm** from the `dplyr::storms` dataset on a world map background.

**Requirements**:

* Choose one storm (e.g., `"Katrina"` in 2005) and filter its rows; order by `date`.
* Plot `lon` vs `lat` as points/line; add a light world map background from `rnaturalearth` (coastlines).
* Animate over `date` using `transition_time(date)` or `transition_reveal(date)` so the path draws over time.

**Stretch** (research & add):

* Add `shadow_mark(alpha = 0.3)` so prior positions remain faint.
* Encode **wind speed** as point size or color and add a titled legend.

```{r}
# Your code here
```

```{r, eval=FALSE}
# ---- Instructor solution (commented) ----
# # World background
# world <- rnaturalearth::ne_countries(scale = "medium", returnclass = "sf")
# library(dplyr)
# data("storms", package = "dplyr")
# trk <- storms %>% filter(name == "Katrina", year == 2005) %>% arrange(date)
#
# p5 <- ggplot() +
#   geom_sf(data = world, fill = "grey95", color = "grey80") +
#   geom_path(data = trk, aes(x = long, y = lat, group = 1), color = "steelblue") +
#   geom_point(data = trk, aes(x = long, y = lat, size = wind, color = wind), alpha = 0.8) +
#   scale_size(range = c(1, 5)) +
#   scale_color_viridis_c(option = "C") +
#   coord_sf(xlim = range(trk$long) + c(-5, 5), ylim = range(trk$lat) + c(-5, 5), expand = FALSE) +
#   labs(title = "Katrina (2005) – {frame_time}", x = NULL, y = NULL, size = "Wind", color = "Wind") +
#   theme_minimal() +
#   transition_time(trk$date) +
#   shadow_mark(alpha = 0.3)
#
# # animate(p5, fps = 18, width = 900, height = 520)
```

---

### Notes & Tips

* `transition_time()` works with continuous time; `transition_states()` with discrete categories; `transition_reveal()` progressively draws along an ordered variable.
* For titles that change over frames, use `{frame_time}`, `{closest_state}`, or `{frame_along}` depending on the transition.
* Use `animate(..., fps = , width = , height = )` to export; save with `anim_save("file.gif")`.
* For long labels beyond the panel, combine `coord_cartesian(clip = "off")` with added right plot margin via `theme(plot.margin = margin(...))`.
