---
title: "In-Class Exercise: ggplot Practice"
author: "Your Name"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
library(ggplot2)
library(dplyr)
library(forcats)
library(palmerpenguins)   # for penguins dataset
```

---

## Exercise 1: Scatterplot with Regression Line  
Dataset: `mpg` (built into ggplot2)  

> Create a scatterplot of engine displacement (`displ`) vs. highway miles per gallon (`hwy`).  
> - Add a **linear regression line** (`geom_smooth(method="lm")`).  
> - Color the points by car class (`class`).  
> What relationship do you see?  


```{r}
ggplot(mpg, aes(x = displ, y = hwy, color = class)) +
  geom_point() +
  geom_smooth(method = "lm")


#You can turn off the confidence band, with se = FALSE
ggplot(mpg, aes(x = displ, y = hwy, color = class)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE)
```

---

## Exercise 2: Bar Chart of Counts  
Dataset: `penguins` (palmerpenguins package)  

> Plot a bar chart showing the number of penguins by `species`.  
> - Reorder the bars so that the most common species comes LAST  

```{r}
penguins |> 
  mutate(species = fct_rev(fct_infreq(species))) |> 
  ggplot(aes(species)) +
  geom_bar()


#or, you could do this:

penguins |> 
  ggplot(aes(fct_rev(fct_infreq(species)))) +
  geom_bar()



# Flip the coordinates so the bars run horizontally. 
penguins |> 
  mutate(species = fct_rev(fct_infreq(species))) |> 
  ggplot(aes(y = species)) +
  geom_bar()

```

---

## Exercise 3: Line Chart of Trends  
Dataset: `economics` (built into ggplot2)  

> Plot the unemployment (`unemploy`) over time (`date`).  
> - Use a line chart.  
> - Add a smoother with `geom_smooth(method="loess")`.  
> What long-term trend do you observe?  

```{r}
ggplot(economics, aes(x = date, y = unemploy)) +
  geom_line() +
  geom_smooth(method = "loess", se = FALSE)
```

---

## Exercise 4: Boxplot with Jittered Points  
Dataset: `diamonds` (built into ggplot2)  

> Create a boxplot of diamond price (`price`) by diamond cut (`cut`).  
> This would mean that price is y, and cut is x
> - Overlay the raw points using `geom_jitter(alpha=0.1)`.  
> - Which cut tends to have the highest median price?  

```{r}
#alpha of 0.1 is the transparency factor
#width of 0.2 is the horizontal width of the data points: +/- 0.2units
ggplot(diamonds, aes(x = cut, y = price)) +
  geom_boxplot() +
  geom_jitter(alpha = 0.1, width = 0.2)
```

---

## Exercise 5: Stacked Bar Plot with Fill  
Dataset: `mpg` (built into ggplot2)  

> Plot a stacked bar chart of number of cars by `class`, filled by `drv` (drive type: f = front-wheel, r = rear-wheel, 4 = 4-wheel).  
> - Use `geom_bar(position="fill")` so the y-axis shows proportions.  
> - Which car class has the highest proportion of 4-wheel drive?  

```{r}
ggplot(mpg, aes(x = class, fill = drv)) +
  geom_bar(position = "fill")
```
