
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

Instructions

This set has 5 problems. Each mirrors an animated example you saw in class and adds one small extension to research.

Use the provided datasets (all are built-in or readily available via popular packages).

Where you see Stretch items, look up the function/argument and implement it.

If you need a hint, check the commented Instructor solution chunk below each problem (keep it commented for the student version).

## Packages

```{r}
# Install (first time only) then load
# install.packages(c("ggplot2","dplyr","gganimate","ggrepel","ggiraph","gapminder","palmerpenguins","sf","rnaturalearth","rnaturalearthdata"))

library(ggplot2)
library(dplyr)
library(gganimate)
library(ggrepel)
library(ggiraph)
library(gapminder)
library(palmerpenguins)
library(sf)
#library(rnaturalearth)
#library(rnaturalearthdata)
```

---

## Problem 1 — Interactive Scatter (penguins)

Create an **interactive scatterplot** of `bill_length_mm` vs `bill_depth_mm` from `palmerpenguins::penguins`.

**Requirements**:

* Color points by `species`.
* Use `geom_point_interactive()` with a meaningful `tooltip` (include species and the two numeric values; format to 1 decimal place).
* Wrap with `girafe()` sized to **6 × 3.7 inches** (golden ratio).
* Add a hover effect so the hovered point is emphasized and the rest fade.

**Stretch** (research & add):

* Add **zoom/pan** controls with `opts_zoom()`.
* Tweak tooltip styling (background light gray; text midnight blue) via `opts_tooltip(css = ...)`.

```{r}
# Your code here
#This filters out all rows with NAs in them
p1 <- penguins |> 
   filter(!is.na(bill_length_mm), !is.na(bill_depth_mm), !is.na(species)) 

#Here, I want to create a new column, which is the row number
# I can use the row number as a "grouping" feature in data_id
#Naturally, if I use row number to group, then ONLY that row is in
#the group. All the other rows are OUTSIDE the group.
#Contrast that, with using "species" for grouping.

#Also, I want a more complicated tooltip. I want to show species, length
#and depth. On new lines.  So, I am using sprintf to format a string
#The \n is a new line
#The %.1f is a "one" decimal place floating point number
p1 <- p1 |> mutate(id = dplyr::row_number(),
                tip = sprintf("%s\nLength: %.1f mm\nDepth: %.1f mm",
                              species, bill_length_mm, bill_depth_mm))

#You have to SAVE the plot into an object, so that we can pass
#it onto girafe
plt <- p1 |>    
  ggplot(aes(bill_length_mm, bill_depth_mm, color = species)) +
   geom_point_interactive(aes(
     tooltip = tip, 
     data_id = id), 
     size = 2.5) +
   theme_minimal()

 girafe(
   ggobj = plt,
   width_svg = 6,
   height_svg = 6 * 0.618,
   options = list(
     opts_hover(css = "opacity:1; stroke:black; stroke-width:1.5px;"),
     opts_hover_inv(css = "opacity:0.2;"),
     opts_tooltip(css = "background:#F5F5F5;color:#191970;padding:4px;border-radius:4px;")
   )
 )
  
```


---



```{r}

# Your code here
# Let us implement the Zoom feature
plt <- p1 |>    
  ggplot(aes(bill_length_mm, bill_depth_mm, color = species)) +
   geom_point_interactive(aes(
     tooltip = tip, 
     data_id = id), 
     size = 2.5) +
   theme_minimal()

 girafe(
   ggobj = plt,
   width_svg = 6,
   height_svg = 6 * 0.618,
   options = list(
     opts_hover(css = "opacity:1; stroke:black; stroke-width:1.5px;"),
     opts_hover_inv(css = "opacity:0.2;"),
     opts_tooltip(css = "background:#F5F5F5;color:#191970;padding:4px;border-radius:4px;"),
     opts_zoom(min = 0.5, max = 5)
   )
 )

```




---

## Problem 2 — Interactive Faceted Plot (mpg)

Create an **interactive faceted scatter** using `ggplot2::mpg` with `hwy` vs `displ`, colored by `class`, faceted by `drv`.

**Requirements**:

* Use `geom_point_interactive()` with a tooltip that shows `manufacturer`, `model`, `hwy`, `displ`.
* Wrap in `girafe()` and remove the legend (use `guides(color = "none")`).


```{r}
# Your code here
 p2 <- mpg |> 
   ggplot(aes(displ, hwy, color = class)) +
   geom_point_interactive(aes(
     tooltip = sprintf("%s %s\nDispl: %.1f\nHwy: %d",
                       manufacturer, model, displ, hwy),
     data_id = class
   ), alpha = 0.8, size = 2) +
   facet_wrap(~ drv) +
   guides(color = "none") +
   theme_minimal()

 girafe(ggobj = p2, 
        width_svg = 6, 
        height_svg = 6*0.618,
        options = list(opts_tooltip(css = "font-size:12px;padding:4px;background:#FAFAFA;border:1px solid #DDDDDD;"),
                       opts_hover(css = "opacity:1; stroke:black; stroke-width:1px;"),
                       opts_hover_inv(css = "opacity:0.2;")))

```
